---
title: "Assignment B4 Option A Exercise 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr) # includes functions required for string manipulation
library(dplyr) # includes case_when() function vectorize multiple if_else statements
library(tibble) # includes tibble() function to create tibbles from scratch
library(testthat) # includes functions required to test and evaluate function behaviors
```

# Go Loco Language : Pig Latin Derivative  

## 1.0 Language Architecture  
**Go Loco** applies a set of explicit rules to translate words, such that each word undergoes a _rearrangement_ and _addition_ transformation. The transformation however will differ slightly depending on the original word structure. 

### 1.1 Rearrangement Rules    
1. **Words that begin with a consonant, followed by a vowel**:

 Move the consonant/vowel pair to the end of the word and then repeat the consonant/vowel pair.  
 Retain the first vowel at the front of the word.  
 
2. **Words that begin with a consonant cluster, followed by a vowel**:

 Move the consonant cluster and following vowel to the end of the word and then repeat this consonant cluster/vowel sound.  
 Retain the first vowel at the front of the word.
 
3. **Words that begin with a vowel**:  

 Move the vowel to the end and with the consonant following it.  
 Repeat this consonant/vowel pair (consonant first) twice.

### 1.2 Addition Rules:
1. Add "loco" to the end of every word.


## 2.0 Examples of Go Loco 

### 2.1 Single Word Examples    
banana -> nanababaloco  
apple -> pleapapaloco

### 2.2 Translating `words` data set in `stringr` package  
The following list is the original `words` data set found in the `stringr` package:

```{r}
words
```

Here is a sample of how the words contained in this data set would be translated into **Go Loco**.

```{r}
# rules applied to words that begin with a consonant/vowel pair
words_trans <-
  tibble("English" = words,
         "Go Loco" = case_when(str_starts(words, "[^aioeu](?=[aiouey])") ~  
                      str_c(str_sub(words, start = 2, end = -1), 
                            str_sub(words, start = 1, end = 2), 
                            str_sub(words, start = 1, end = 2), 
                            "loco"),
# rules applied to words that begin with a consonant cluster/vowel sound
                    str_starts(words, "[^aioeu](?![aiouey])") ~ 
                      str_c(str_extract(words, "([aioeuy]).*"), 
                            str_extract(words, "^[^aioue]+[aiouey]+"),
                            str_extract(words, "^[^aioue]+[aiouey]+"),
                            "loco"),
# rules applied to words that begin with a vowel
                    str_starts(words, "[aioeu]") ~
                      str_c(str_extract(words, "(?<=[^aioue]).*"), 
                            str_extract(words, "(^[aioue]+[^aioue])"),
                            str_extract(words, "(^[aioue]+[^aioue])"),
                            str_extract(words, "(^[aioue])"),
                            "loco")))

words_trans
```

Note that the word must be at least 2 letters long to be translated into **Go Loco** e.g. "a" does not have a **Go Loco** translation.

## 3.0 Function `goloco`  

The following code creates a function called `goloco` that takes a string or character vector input for `words` and outputs a character vector corresponding to the **Go Loco** translation of the input. 

```{r} 
#' @title Translate word(s) into Go Loco "Pig Latin" derivative language
#'
#' @param words A string or character vector to translate into Go Loco
#'
#' @return A character vector with the translated input
#' @import stringr
#' @import dplyr
#' @export
#'
#' @examples
#' goloco(banana)
#' goloco(c("pineapple", "melon", "apple"))
#' goloco(str_split_1("the quick fox jumped over the lazy moon", " "))
#' 
goloco <- function(words) {
  if(is.numeric(words)) {
        stop("Sorry! You can't make numbers go loco :(")
    }
# rules applied to words that begin with a consonant/vowel pair
  case_when(str_starts(words, "[^aioeu](?=[aiouey])") ~  
                      str_c(str_sub(words, start = 2, end = -1), 
                            str_sub(words, start = 1, end = 2), 
                            str_sub(words, start = 1, end = 2), 
                            "loco"),
# rules applied to words that begin with a consonant cluster/vowel sound
                    str_starts(words, "[^aioeu](?![aiouey])") ~ 
                      str_c(str_extract(words, "([aioeuy]).*"), 
                            str_extract(words, "^[^aioue]+[aiouey]+"),
                            str_extract(words, "^[^aioue]+[aiouey]+"),
                            "loco"),
# rules applied to words that begin with a vowel
                    str_starts(words, "[aioeu]") ~
                      str_c(str_extract(words, "(?<=[^aioue]).*"), 
                            str_extract(words, "(^[aioue]+[^aioue])"),
                            str_extract(words, "(^[aioue]+[^aioue])"),
                            str_extract(words, "(^[aioue])"),
                            "loco"))
}
```

## 4.0 Examples of `goloco`

### 4.1 Translating single words using `goloco`

Here is an example of applying `goloco` to translate single words. 

```{r}
goloco("banana")
goloco("apple")
goloco("pineapple")
```

### 4.2 Translating a sentence using `goloco`

Here is a example of applying `goloco` to translate a sentence.

```{r}
# requires splitting the sentence into its component words, using the separator " " 
goloco(str_split_1("we all like to eat apples and bananas", " "))
```

### 4.3 Translating `words` data set in `stringr` package using `goloco`  

Recall the `words` data set found in the `stringr` package:

```{r}
head(words, n = 100)
```

Here is an example of applying `goloco` to translate the character vector `words`.

```{r}
goloco(stringr::words)
```

## 5.0 Tests for `goloco` Functionality  

The following section formally tests the `goloco` function given various types of input to test the robustness of its intended use and its response to inappropriate inputs. 

### 5.1 String input, starts with consonant 
This test evaluates if `goloco` correctly translates a string that begins with a consonant. 

```{r}
test_that("translating words that starts with consonant", {
  expect_identical(object = goloco("lollipop"),
                   expected = "ollipoplololoco")
})
```

### 5.2 String input, starts with vowel 
This test evaluates if `goloco` correctly translates a string that begins with a vowel. 

```{r}
test_that("translating words that starts with vowel", {
  expect_identical(object = goloco("alligator"),
                   expected = "ligatoralalaloco")
})
```

### 5.3 Character vector input 
This test evaluates if `goloco` correctly translates a vector of character strings.

```{r}
test_that("translating vector of character strings", {
  expect_identical(object = goloco(c("balloon", "clown", "cake")),
                   expected = c("alloonbabaloco", "ownclocloloco", "akecacaloco"))
})
```

### 5.4 Numeric input
This test evaluates if `goloco` correctly prevents numeric inputs. 

```{r}
test_that("error given numeric input", {
  expect_error(object = goloco(123))
})
```

