---
title: "Assignment B4 Option A Exercise 1"
author: "Renessa"
date: "2023-12-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Sources of data
library(gutenbergr) # includes functions required to load texts from Project Gutenberg

# Data manipulation
library(tidytext) # includes functions required to tidy text data
library(dplyr) # includes functions required for data wrangling
library(stringr) # includes functions required for string manipulation and detection
library(forcats) # includes functions to manipulate factors and levels in data

# Data visualization
library(ggplot2) # includes functions required for data visualization
```

# 1.0 A Christmas Carol by Charles Dickens

The following code retrieves "A Christmas Carol" by Charles Dickens from the Project Gutenberg database. 

```{r}
xmascarol_crude <- gutenberg_download(19337)
print(xmascarol_crude)
```

## 1.1 Cleaning the data for text mining  

Here is a tibble of the text with line and chapter denotations. 

```{r}
xmascarol_df <- xmascarol_crude %>% 
  select(text) %>% 
  mutate(line = row_number(),
         chapter = cumsum(str_count(text, "STAVE(?=\\s[A-Z]+)")))
print(xmascarol_df)
```

In order to determine the most commonly used words, the above tibble will need to be restructured to only contain one word per row, and standardize the caps lock to ensure words at the beginning vs middle of a sentence correspond to a single identifier. 

```{r}
xmascarol_words <- xmascarol_df %>% unnest_tokens(word, text) 
print(xmascarol_words)
```

Next, "stopwords" will need to be removed to determine which meaningful words appear most often in the text. Stopwords were referenced from `tidytext::stop_words`, including stopwords from three lexicons. 

```{r}
xmascarol_words_clean <- xmascarol_words %>% anti_join(stop_words, by = "word")
print(xmascarol_words_clean)
```

## 1.2 Determining the most common words used  

Lastly, the above tibble can be manipulated to output a tibble of words sorted by frequency at which they appear. 

```{r}
xmascarol_words_clean_n <- 
  xmascarol_words_clean %>% 
  group_by(word) %>% 
  summarize(frequency = n()) %>% 
  arrange(desc(frequency))

print(xmascarol_words_clean_n)
```

## 1.3 Plotting the most common words used  

Here is a bar chart of the 30 most commonly used words in _A Christmas Carol_ by Charles Dickens. 

```{r}
ggplot(data = xmascarol_words_clean_n[1:30, ],
       aes(x = reorder(word, frequency), # ordering the bars according to frequency
           y = frequency)) +
  geom_col() +
  geom_text(aes(label = frequency),
             size = 2.5,
             nudge_y = 5) +
  labs(x = "Word",
       y = "Frequency") +
  coord_flip() +
  theme_bw()
```

